---
title: "R Notebook"
output: html_notebook
---

# Set up and load relevant packages

```{r setup}
if (!require("pacman")) install.packages("pacman"); 
library(pacman)
pacman::p_load(tidyverse, dplyr, nycflights13)

#pacman installs a package before loading it if it is not already installed.
```

Make a scatterplot of hwy vs cyl.

```{r}
ggplot(mpg, aes(x = hwy, y = cyl)) +
  geom_point()
```

What happens if you make a scatterplot of class vs drv? Why is the plot not useful? What happens if you use geom_count() instead?

```{r}
ggplot(mpg, aes(x = class, y = drv)) +
  geom_point()

ggplot(mpg, aes(x = class, y = drv)) +
  geom_count()
```

What’s gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

```{r}
#The argument colour = "blue" is included within the mapping argument, and as such, it is treated as an aesthetic, which is a mapping between a variable and a value. In the expression, colour = "blue", "blue" is interpreted as a categorical variable which only takes a single value "blue".

#The following code produces the expected result.
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), colour = "blue")
```


Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r}
ggplot(mpg, aes(x = displ, y = hwy, colour = cty)) +
  geom_point()
```

What happens if you map the same variable to multiple aesthetics?

```{r}
ggplot(mpg, aes(x = displ, y = hwy, colour = hwy, size = displ)) +
  geom_point()
```

```{r}
#The code works and produces a plot, even if it is a bad one. Mapping a single variable to multiple aesthetics is redundant. Because it is redundant information, in most cases avoid mapping a single variable to multiple aesthetics.
```


What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

```{r}
# Stroke changes the size of the border for shapes (21-25). These are filled shapes in which the color and size of the border can differ from that of the filled interior of the shape.

ggplot(mtcars, aes(wt, mpg)) +
  geom_point(shape = 21, colour = "black", fill = "white", size = 5, stroke = 5)

```

What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.

```{r}
ggplot(mpg, aes(x = displ, y = hwy, colour = displ < 5)) +
  geom_point()
  
# Aesthetics can also be mapped to expressions like displ < 5. The ggplot() function behaves as if a temporary variable was added to the data with values equal to the result of the expression. In this case, the result of displ < 5 is a logical variable which takes values of TRUE or FALSE.

```


What plots does the following code make? What does . do?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(drv ~ .)

#The symbol . ignores that dimension when faceting. For example, drv ~ . facet by values of drv on the y-axis.
```



Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
geom_point() + 
geom_smooth(se = FALSE)

#This code produces a scatter plot with displ on the x-axis, hwy on the y-axis, and the points colored by drv. There will be a smooth line, without standard errors, fit through each drv group.
```

What does the se argument to geom_smooth() do?

```{r}
#It adds standard error bands to the lines. By default se = TRUE

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, colour = drv)) +
  geom_point() +
  geom_smooth(se = TRUE)

```

What does geom_col() do? How is it different to geom_bar()?

```{r}
#The geom_col() function has different default stat than geom_bar(). The default stat of geom_col() is stat_identity(), which leaves the data as is. The geom_col() function expects that the data contains x values and y values which represent the bar height.

#The default stat of geom_bar() is stat_count(). The geom_bar() function only expects an x variable. The stat, stat_count(), preprocesses input data by counting the number of observations for each value of x. The y aesthetic uses the values of these counts.
```

Create one plot on the fuel economy data with customized title, subtitle, caption, x, y, and colour labels.

```{r}
ggplot(
  data = mpg,
  mapping = aes(x = fct_reorder(class, hwy), y = hwy)
) +
  geom_boxplot() +
  coord_flip() +
  labs(
    title = "Compact Cars have > 10 Hwy MPG than Pickup Trucks",
    subtitle = "Comparing the median highway mpg in each class",
    caption = "Data from fueleconomy.gov",
    x = "Car Class",
    y = "Highway Miles per Gallon"

```

